---
title: "Cancer Screening 2021"
author: "David Fong"
date: "25th May 2021"
output:
  rmdformats::readthedown:
    toc_depth: 4
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(dplyr)
library(magrittr)
library(ggplot2)
```

'Low-frequency' mental health conditions : schizophrenia, bipolar disorder

Of the ninety-four (94) active patients at Kensington who EITHER have a coded
diagnosis of a low-frequency mental health disorder OR have a major tranquilliser
in their medication list, nineteen (19) do NOT have a coded diagnosis of a
low-frequency mental health disorder.

## Bowel cancer screening

Age 50 to 74 years inclusive, active patients.


```{r data_bowel}
fobt <- data.frame(
  clinic = character(),
  lowfreq = logical(),
  status = character(),
  n = integer()
)

fobt <- fobt %>%
  add_row(clinic = "Kensington", lowfreq = TRUE, status = "UTD", n = 8) %>%
  add_row(clinic = "Kensington", lowfreq = TRUE, status = "Late", n = 2+3+9) %>%
  # the three numbers in 'Late' are actually 2-3, 3-4 and >4 years
  add_row(clinic = "Kensington", lowfreq = TRUE, status = "Never", n = 62) %>%
  add_row(clinic = "Kensington", lowfreq = FALSE, status = "UTD", n = 204) %>%
  add_row(clinic = "Kensington", lowfreq = FALSE, status = "Late", n = 61+22+85) %>%
  add_row(clinic = "Kensington", lowfreq = FALSE, status = "Never", n = 610) %>%
  add_row(clinic = "Paisley", lowfreq = TRUE, status = "UTD", n = 15) %>%
  add_row(clinic = "Paisley", lowfreq = TRUE, status = "Late", n = 5+4+7) %>%
  # the three numbers in 'Late' are actually 2-3, 3-4 and >4 years
  add_row(clinic = "Paisley", lowfreq = TRUE, status = "Never", n = 62) %>%
  add_row(clinic = "Paisley", lowfreq = FALSE, status = "UTD", n = 306) %>%
  add_row(clinic = "Paisley", lowfreq = FALSE, status = "Late", n = 106+46+126) %>%
  add_row(clinic = "Paisley", lowfreq = FALSE, status = "Never", n = 857) %>%
  add_row(clinic = "Collingwood", lowfreq = TRUE, status = "UTD", n = 15) %>%
  add_row(clinic = "Collingwood", lowfreq = TRUE, status = "Late", n = 3+7+8) %>%
  # the three numbers in 'Late' are actually 2-3, 3-4 and >4 years
  add_row(clinic = "Collingwood", lowfreq = TRUE, status = "Never", n = 92) %>%
  add_row(clinic = "Collingwood", lowfreq = FALSE, status = "UTD", n = 282) %>%
  add_row(clinic = "Collingwood", lowfreq = FALSE, status = "Late", n = 74+51+120) %>%
  add_row(clinic = "Collingwood", lowfreq = FALSE, status = "Never", n = 1338) %>%
  add_row(clinic = "Fitzroy", lowfreq = TRUE, status = "UTD", n = 15) %>%
  add_row(clinic = "Fitzroy", lowfreq = TRUE, status = "Late", n = 3+4+7) %>%
  # the three numbers in 'Late' are actually 2-3, 3-4 and >4 years
  add_row(clinic = "Fitzroy", lowfreq = TRUE, status = "Never", n = 104) %>%
  add_row(clinic = "Fitzroy", lowfreq = FALSE, status = "UTD", n = 208) %>%
  add_row(clinic = "Fitzroy", lowfreq = FALSE, status = "Late", n = 62+39+78) %>%
  add_row(clinic = "Fitzroy", lowfreq = FALSE, status = "Never", n = 1042)
```

### Plots

```{r plot_bowel_n}
ggplot(
  fobt %>%
    mutate(clinic_group = paste(clinic, lowfreq)),
  aes(
    fill = factor(status, levels = c("UTD", "Late", "Never")),
    y = n, x = clinic_group
  )) +
  geom_bar(position = "stack", stat = "identity") +
  coord_flip() + 
  ggtitle("Number of patients with up-to-date bowel cancer screening") +
  ylab("n") + xlab("Clinic and low-frequency mental health condition status") +
  guides(fill = guide_legend("Screening status"))

```

```{r plot_bowel_proportion}
ggplot(
  fobt %>%
    mutate(clinic_group = paste(clinic, lowfreq)),
  aes(
    fill = factor(status, levels = c("UTD", "Late", "Never")),
    y = n, x = clinic_group
  )) +
  geom_bar(position = "fill", stat = "identity") +
  coord_flip() + 
  ggtitle("Proportion of patients with up-to-date bowel cancer screening") +
  ylab("Proportion") + xlab("Clinic and low-frequency mental health condition status") +
  guides(fill = guide_legend("Screening status"))

```

## Cervical cancer screening

Age 30 to 74 years inclusive, active patients

```{r data_cst}
cst <- data.frame(
  clinic = character(),
  lowfreq = logical(),
  status = character(),
  n = integer()
)

cst <- cst %>%
  add_row(clinic = "Kensington", lowfreq = TRUE, status = "UTD", n = 36) %>%
  add_row(clinic = "Kensington", lowfreq = TRUE, status = "Late", n = 11) %>%
  add_row(clinic = "Kensington", lowfreq = TRUE, status = "Never", n = 14) %>%
  add_row(clinic = "Kensington", lowfreq = FALSE, status = "UTD", n = 692) %>%
  add_row(clinic = "Kensington", lowfreq = FALSE, status = "Late", n = 117) %>%
  add_row(clinic = "Kensington", lowfreq = FALSE, status = "Never", n = 254) %>%
  add_row(clinic = "Paisley", lowfreq = TRUE, status = "UTD", n = 49) %>%
  add_row(clinic = "Paisley", lowfreq = TRUE, status = "Late", n = 14) %>%
  add_row(clinic = "Paisley", lowfreq = TRUE, status = "Never", n = 23) %>%
  add_row(clinic = "Paisley", lowfreq = FALSE, status = "UTD", n = 1043) %>%
  add_row(clinic = "Paisley", lowfreq = FALSE, status = "Late", n = 172) %>%
  add_row(clinic = "Paisley", lowfreq = FALSE, status = "Never", n = 517) %>%
  add_row(clinic = "Collingwood", lowfreq = TRUE, status = "UTD", n = 47) %>%
  add_row(clinic = "Collingwood", lowfreq = TRUE, status = "Late", n = 19) %>%
  add_row(clinic = "Collingwood", lowfreq = TRUE, status = "Never", n = 33) %>%
  add_row(clinic = "Collingwood", lowfreq = FALSE, status = "UTD", n = 840) %>%
  add_row(clinic = "Collingwood", lowfreq = FALSE, status = "Late", n = 218) %>%
  add_row(clinic = "Collingwood", lowfreq = FALSE, status = "Never", n = 1399) %>%
  add_row(clinic = "Fitzroy", lowfreq = TRUE, status = "UTD", n = 49) %>%
  add_row(clinic = "Fitzroy", lowfreq = TRUE, status = "Late", n = 19) %>%
  add_row(clinic = "Fitzroy", lowfreq = TRUE, status = "Never", n = 32) %>%
  add_row(clinic = "Fitzroy", lowfreq = FALSE, status = "UTD", n = 673) %>%
  add_row(clinic = "Fitzroy", lowfreq = FALSE, status = "Late", n = 183) %>%
  add_row(clinic = "Fitzroy", lowfreq = FALSE, status = "Never", n = 597)
```

### Plots

```{r plot_cst_n}
ggplot(
  cst %>%
    mutate(clinic_group = paste(clinic, lowfreq)),
  aes(
    fill = factor(status, levels = c("UTD", "Late", "Never")),
    y = n, x = clinic_group
  )) +
  geom_bar(position = "stack", stat = "identity") +
  coord_flip() + 
  ggtitle("Number of patients with up-to-date cervical cancer screening") +
  ylab("n") + xlab("Clinic and low-frequency mental health condition status") +
  guides(fill = guide_legend("Screening status"))

```

```{r plot_cst_proportion}
ggplot(
  cst %>%
    mutate(clinic_group = paste(clinic, lowfreq)),
  aes(
    fill = factor(status, levels = c("UTD", "Late", "Never")),
    y = n, x = clinic_group
  )) +
  geom_bar(position = "fill", stat = "identity") +
  coord_flip() + 
  ggtitle("Proportion of patients with up-to-date cervical cancer screening") +
  ylab("Proportion") + xlab("Clinic and low-frequency mental health condition status") +
  guides(fill = guide_legend("Screening status"))

```