---
title: "COVID vaccination effectiveness"
author: "David Fong"
date: "26th May 2021"
output:
  rmdformats::readthedown:
    toc_depth: 4
    code_folding: hide
references:
- id: ourworldindata
  title: "Coronavirus Pandemic (COVID-19)"
  author:
  - Max Roser
  - Hannah Ritchie
  - Esteban Ortiz-Ospina
  - Joe Hasell
  issued:
      year: 2020
  URL: https://ourworldindata.org/coronavirus
- id: ourworldindatavax
  title: "A global database of COVID-19 vaccinations"
  author:
  - Ritchie, Mathieu, E.
  - Ritchie, H.
  - Ortiz-Ospina, E. et al.
  container-title: Nat Hum Behav
  issued:
    year: 2021
  URL: https://doi.org/10.1038/s41562-021-01122-8
- id: hungarytoday
  title: "Hungary’s Vaccinations: Pfizer Used for Most, Sinopharm Catching Up"
  author:
  - Tamás Vaski
  issued:
    year: 2021
  container-title: Hungary Today
  URL: https://hungarytoday.hu/hungary-vaccinations-coronavirus-operative-board-statistics-vaccination-amounts-number-of-inoculated/
- id: ecdc
  title: "COVID-19 Vaccine rollout overview (Week 19, 2021)"
  container-title: European Center for Disease Prevention and Control
  issued:
    year: 2021
  URL: https://covid19-vaccine-report.ecdc.europa.eu/#3_Uptake_of_at_least_one_vaccine_dose_among_adults
- id: ctvnews
  title: "Risks of rare blood clot even lower with second dose of AstraZeneca, experts say"
  container-title: "CTV News"
  issued:
    year: 2021
  URL: https://www.ctvnews.ca/health/coronavirus/risks-of-rare-blood-clot-even-lower-with-second-dose-of-astrazeneca-experts-say-1.5437073
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE}
library(tidyverse)
library(dplyr)
library(magrittr)
library(lubridate)
library(kableExtra)
library(ggplot2)
library(RColorBrewer)
```

## Objective

Is Astra-Zeneca doses administered (at least a single dose) - as a proportion of
*all* COVID vaccination administered - related to a national reduction in
COVID-19 cases and death?

## Data

Countries selected on the basis of overall relatively high rate of overall coverage (40%
as of May 2021, including vaccines which are *not* Astra-Zeneca), the availability of
COVID case and death data approximately three weeks after 35% single-dose COVID vaccination
coverage achieved and high rates of COVID-19 cases and death in a December 2020
to January 2021 peak.

Both Hungary and Canada experienced additional 'peaks' between January 2021 and
the time they achieved 35% coverage.

Canada achieved 35% single-dose coverage on 5th May 2021.
Germany achieved 35% single-dose coverage on 12th May 2021.
(As of 25th May 2021, it is not quite yet three weeks since either of those two dates)

Daily number of cases (seven day rolling average), daily number of deaths (seven day rolling average) and proportion
of population vaccinated derived from [Our World in Data](https://github.com/owid/covid-19-data/tree/master/public/data/).

### Timepoints

* Day of maximum cases in December 2020 to January 2021
* Day of maximum deaths in December 2020 to January 2021
* Day of 35% single-dose COVID vaccination coverage (any COVID-19 vaccine)
* Three weeks after the day of 35% single-dose COVID vaccination coverage
  + vaccines are considered effective one to two weeks after delivery of dose
  + COVID deaths tend to lag case-finding by one to two weeks

```{r read_data, message = FALSE}
# from "Our World in Data"
owid = read_csv(
  "https://covid.ourworldindata.org/data/owid-covid-data.csv",
  col_names=TRUE,
  col_types="ccc?nnnnnnnnnnnnnnnnnnnnnnnnnnnnncnnnnnnnnnnnnnnnnnnnnnnnnn",
  skip_empty_rows = FALSE
)

# European Center for Disease Prevention and Control, Week 19 of 2021
# used to derive use of different vaccine brands in Hungary
hungary_vax <- read_csv("~/Rprojects/coHealthKensington/Hungary_COVIDvax_dataset_2021-W19.csv")
germany_vax <- read_csv("~/Rprojects/coHealthKensington/Germany_COVIDvax_dataset_2021-W19.csv")
```

```{r data}
countries <- owid %>%
  select(location, date, new_cases_smoothed, new_deaths_smoothed, people_vaccinated_per_hundred) %>%
  filter(
    location == "Canada" | location == "Israel" |
      location == "United States" | location == "United Kingdom" |
      location == "Hungary" | location == "Germany")

# Iceland only had approximately 10 cases per month in Dec-January 2021
# (and even less cases at the time of 35% vaccination)
# Finland had less than 10 deaths per months in Dec-January 2021
# Cyprus reached 35% vaccination 14th May 2021, and had
# less than 10 deaths per day in Dec-January 2021
# Unable to find data on Uruguay vaccine brand numbers,
# though in Uruguay peak deaths in Dec-Jan 2021 was also approximately 10/day

cases <- countries[FALSE,] # 'empty' version of countries data

for (i in unique(countries$location)) {
  country <- countries %>% filter(location == i)
  
  # days of max cases and deaths in January 2021
  max_cases <- country %>%
    filter(date >= ymd("2020-12-01") & date <= ymd("2021-01-31")) %>%
    slice_max(new_cases_smoothed)
  max_deaths <- country %>%
    filter(date >= ymd("2020-12-01") & date <= ymd("2021-01-31")) %>%
    slice_max(new_deaths_smoothed)
  
  # day of 35% single-dose vaccination coverage, and three weeks after that
  vax_35_date <- country %>%
    filter(people_vaccinated_per_hundred >= 35 & people_vaccinated_per_hundred < 40) %>%
    slice_min(people_vaccinated_per_hundred)
  post35_3week <- country %>%
    filter(date == nth(date, which.min(abs(date - (vax_35_date$date + weeks(3))))))
  
  cases <- cases %>% 
    add_row(as.data.frame(max_cases)) %>%
    add_row(as.data.frame(max_deaths)) %>%
    add_row(as.data.frame(vax_35_date)) %>%
    add_row(as.data.frame(post35_3week))
}

cases <- cases %>%
  rename(
    country = location,
    cases = new_cases_smoothed, 
    # 7-day rolling average
    deaths = new_deaths_smoothed, 
    # 7-day rolling average
    vaccinated = people_vaccinated_per_hundred 
    # Total number of people who received at least one vaccine dose per 100 people in the total population
  )
```

```{r}
cases %>%
  kbl(
    caption = "January peak cases/deaths (7-day smoothed) and 3-weeks after 35% single-dose coverage",
    col.names = c("Country", "Date", "Daily cases", "Daily deaths", "Vaccinated (at least one dose) per hundred"),
    digits = 1
  ) %>%
  kable_styling()
```

## Reduction in cases and deaths after 35% of single-dose coverage achieved

Calculate ratio of reduction in cases and deaths (seven day rolling average) three
weeks after single-dose coverage is achieved.

```{r}
effectiveness <- data.frame(
  country = character(),
  max_cases = numeric(),
  max_deaths = numeric(),
  post35_3wk = ymd(),
  post35_3wk_cases = numeric(),
  post35_3wk_deaths = numeric(),
  astrazeneca_proportion = numeric()
)

for (i in unique(cases$country)) {
  country_cases <- cases %>% filter(country == i)
  peak_cases <- country_cases %>%
    filter(date >= ymd("2020-12-01") & date <= ymd("2021-01-31")) %>%
    filter(cases == max(cases)) %>%
    pull(cases)
  peak_deaths <- country_cases %>%
    filter(date >= ymd("2020-12-01") & date <= ymd("2021-01-31")) %>%
    filter(deaths == max(deaths)) %>%
    pull(deaths)
  vax_35_date <- country_cases %>%
    filter(vaccinated >= 35 & vaccinated < 36) %>%
    pull(date) # date of vaccinations between 0.35 and 0.36
  post35_3week <- country_cases[which.min(abs((vax_35_date + weeks(3)) - country_cases$date)),] 
  # closest date to 3 weeks post vax_35_date
  
  effectiveness <- effectiveness %>%
    add_row(
      country = i,
      max_cases = peak_cases,
      max_deaths = peak_deaths,
      post35_3wk = post35_3week$date,
      post35_3wk_cases = post35_3week$cases,
      post35_3wk_deaths = post35_3week$deaths
      )
}

effectiveness <- effectiveness %>%
  mutate(
    case_reduction_ratio = max_cases/post35_3wk_cases,
    death_reduction_ratio = max_deaths/post35_3wk_deaths
  ) %>%
  mutate(
    log_case_reduction = log(case_reduction_ratio),
    log_death_reduction = log(death_reduction_ratio)
  )
```

```{r}
effectiveness %>%
  select(country, case_reduction_ratio, death_reduction_ratio) %>%
  kbl(
    caption = "Reduction in COVID-19 cases and deaths after 35% single-dose vaccination coverage",
    col.names = c("Country", "Case reduction ratio", "Death reduction ratio"),
    digits = 2
  ) %>%
  kable_styling()
```

## Astra-Zeneca usage

Neither Israel or United States used Astra-Zeneca vaccines at all.

Hungary and German vaccination brand delivery available via the European Centre for Disease and Control. 

Vaccines codes (ECDC):

* COM - Pfizer/Comirnaty
* BECNBG - CNBG/Sinopharm
* SPU - Sputnik V
* AZ - Astra-Zeneca
* MOD - Moderna
* JANSS - Janssen

```{r astrazeneca_usage, message=FALSE}
effectiveness[effectiveness$country == "United States", "astrazeneca_proportion"] <- 0
effectiveness[effectiveness$country == "Israel", "astrazeneca_proportion"] <- 0

# sourced from European Centre for Disease Prevention and Control
# week 19 of 2021
hungary_summary <- hungary_vax %>%
  select(`Vaccine brand`, `First dose`) %>%
  rename(Vaccine = `Vaccine brand`, Doses = `First dose`) %>%
  group_by(Vaccine) %>%
  summarize(Doses = sum(Doses))
# unlike most other European countries, Hungary sourced more than a third
# of its COVID vaccines from Russian or Chinese brands

effectiveness[effectiveness$country == "Hungary", "astrazeneca_proportion"] <- 
  hungary_summary[hungary_summary$Vaccine == "AZ", "Doses"]/sum(hungary_summary$Doses)

hungary_summary %>%
  arrange(desc(Doses)) %>%
  kbl(caption = "Hungary cumulative first-dose administered COVID vaccine brands, week 19 of 2021") %>%
  kable_styling()
```

```{r message=FALSE}
# sourced from European Centre for Disease Prevention and Control
# week 19 of 2021
germany_summary <- germany_vax %>%
  select(`Vaccine brand`, `First dose`) %>%
  rename(Vaccine = `Vaccine brand`, Doses = `First dose`) %>%
  group_by(Vaccine) %>%
  summarize(Doses = sum(Doses))

effectiveness[effectiveness$country == "Germany", "astrazeneca_proportion"] <- 
  germany_summary[germany_summary$Vaccine == "AZ", "Doses"]/sum(germany_summary$Doses)

germany_summary %>%
  arrange(desc(Doses)) %>%
  kbl(caption = "Germany cumulative first-dose administered COVID vaccine brands, week 19 of 2021") %>%
  kable_styling()
```

[A CTV news report "Risks of rare blood clot even lower with second dose of AstraZeneca, experts say"](https://www.ctvnews.ca/health/coronavirus/risks-of-rare-blood-clot-even-lower-with-second-dose-of-astrazeneca-experts-say-1.5437073), dated 20th May 2021, reports:

* *'Britain, which has had the AstraZeneca shot in circulation much longer than Canada, has administered more than 23 million first doses of the vaccine'*
* *'In Canada, approximately 2.1 million people have received one dose of the AstraZeneca vaccine'*

```{r}
# CTV news report dated May 20, 2021
# 'Britain, which has had the AstraZeneca shot in circulation much longer than Canada, has administered more than 23 million first doses of the vaccine'
# 'In Canada, approximately 2.1 million people have received one dose of the AstraZeneca vaccine'

effectiveness[effectiveness$country == "United Kingdom", "astrazeneca_proportion"] <-
  23000000/(owid %>%
              filter(location == "United Kingdom", date == as.Date("2021-05-20")) %>%
              pull(people_vaccinated))

effectiveness[effectiveness$country == "Canada", "astrazeneca_proportion"] <-
  2100000/(owid %>%
              filter(location == "Canada", date == as.Date("2021-05-20")) %>%
              pull(people_vaccinated))

owid %>%
  filter(location == "United Kingdom" | location == "Canada") %>%
  filter(date == as.Date("2021-05-20")) %>%
  select(location, people_vaccinated) %>%
  rename(Country = location, `Total vaccinated` = people_vaccinated) %>%
  kbl(caption = "Total vaccinated, as of 20th May 2021 (source 'Our World in Data')") %>%
  kable_styling()

```

## Relationship between AstraZeneca vaccine usage and reduction in COVID-19 cases and deaths

Comparing the rates of cases and deaths from peak of December 2020-January 2021 to three weeks
after 35% single-dose coverage has been achieved.

```{r}
ggplot(
  effectiveness,
  aes(
    x = astrazeneca_proportion, y = case_reduction_ratio,
    label = country
  )) +
  xlim(-0.05, .8) +
  geom_point() +
  geom_text(aes(label = country), hjust = -0.1, vjust = 0) +
  ggtitle("Reduction in COVID-19 cases compared to January 2021 peak", subtitle = "Three weeks after 35% single-dose coverage") +
  xlab("Astra-Zeneca as proportion of administered vaccines") +
  ylab("Ratio of case reduction")
  
```

```{r}
ggplot(
  effectiveness,
  aes(
    x = astrazeneca_proportion, y = log_death_reduction,
    label = country
  )) +
  xlim(-0.05, .8) +
  geom_point() +
  geom_text(aes(label = country), hjust = -0.1, vjust = 0) +
  ggtitle("Reduction in COVID-19 deaths compared to January 2021 peak", subtitle = "Three weeks after 35% single-dose coverage") +
  xlab("Astra-Zeneca as proportion of administered vaccines") +
  ylab("Ratio of death reduction")
```

## References